---
title: "Comorbidity ICD Analysis"
author: "Daniel Wray 105134386"
date: "3/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This document investigates specific ICDs and creates two boolian variables that identify patients with ICDs indicative of Diabetes/DKA and Cancer/Oncology/Chemotherapy.

### Reading and Organizing the Data

This document requires the filtered csv's with non-target observations removed!

```{r}
library(readr)
library(dplyr)
core2019 <- read_csv("Core_2019_Filtered.csv")
core2018 <- read_csv("Core_2018_Filtered.csv")
core2017 <- read_csv("Core_2017_Filtered - Core_2017_Filtered.csv.csv")

core <- bind_rows(core2017, core2018, core2019)

### Extracting only the ICD codes from the dataset ###
cm_data <- core[,c(1, 20:59)]
cm_data_clean <- core[,c(20:59)]

```


### INVESTIGATING COMORBIDITIES

For the first three characters, for codes like E10

```{r}
#### Investigation of first three characters of ICD codes ####
### Only extracting the first three characters of the ICD codes
simplified_3 <- substr(as.matrix(cm_data_clean), 1, 3)

## Table of first 3 characters of codes ##
table(simplified_3)

## Creating dataset with observations marked if they contain an ICD code with certain first 3 characters
cm_simplified_3 <- names(table(simplified_3))
cm_type_data_3 <- data.frame(matrix(ncol = 317, nrow = 68))
for(i in 1:317){
  cm_type_data_3[,i] <- apply(simplified_3, 1, function(r) any(r %in% cm_simplified_3[i]))
}
colnames(cm_type_data_3) <- cm_simplified_3

## Totals of all codes
colSums(cm_type_data_3)

```

For the first three characters, for codes like C

```{r}
#### Investigation of first character of ICD codes ####
### Only extracting the first character of the ICD codes
simplified_1 <- substr(as.matrix(cm_data_clean), 1, 1)

## Table of first character of codes
table(simplified_1)

## Creating dataset with observations marked if they contain an ICD code that starts with a certain character 
cm_simplified_1 <- names(table(simplified_1))
cm_type_data_1 <- data.frame(matrix(ncol = 23, nrow = 68))
for(i in 1:23){
  cm_type_data_1[,i] <- apply(simplified_1, 1, function(r) any(r %in% cm_simplified_1[i]))
}
colnames(cm_type_data_1) <- cm_simplified_1

## Totals of all codes
colSums(cm_type_data_1)


### Looking at specific ICD code distributions
Es <- as.matrix(cm_data_clean)[which(simplified_1 == "E")]
table(Es)

Cs <- as.matrix(cm_data_clean)[which(simplified_1 == "C")]
table(Cs)

Zs <- as.matrix(cm_data_clean)[which(simplified_1 == "Z")]
table(Zs)
```



### Investigating Distributions and Creating Bools

```{r}
## First diabetes
DIABETES <- vector(length = 68)

DIABETES[which(cm_type_data_3$E10 == TRUE)] <- TRUE
DIABETES[which(cm_type_data_3$E11 == TRUE)] <- TRUE
DIABETES[which(cm_type_data_3$E12 == TRUE)] <- TRUE
DIABETES[which(cm_type_data_3$E13 == TRUE)] <- TRUE
DIABETES[which(cm_type_data_3$E14 == TRUE)] <- TRUE

table(DIABETES)


## Then Cancer
CANCER <- vector(length = 68)

CANCER[which(cm_type_data_1$C == TRUE)] <- TRUE

table(CANCER)


## Finally immunosuppressants
IMMUNO <- vector(length = 68)

IMMUNO[which(cm_type_data_3$Z92 == TRUE)] <- TRUE

table(IMMUNO)


#### Creating Tables for Analysis ####

data_for_tables <- data.frame(core, DIABETES, CANCER, IMMUNO)
table(data_for_tables[c("DIABETES", "DIED")])
table(data_for_tables[c("CANCER", "DIED")])
table(data_for_tables[c("IMMUNO", "DIED")])
table(data_for_tables[c("DIABETES", "CANCER", "DIED")])

```


### Dataframe Creation
```{r}
ICD_two_bool_variables <- data.frame(DIABETES, CANCER)
write.csv(ICD_two_bool_variables, "ICD Diabetes and Cancer.csv")
```
