title: Number of Procedures and CANCER + DIABETES over COMOR Analysis
author: Original code by {Matthew Mallory}, edited by {Daniel Wray}
date: {3/9/2022}
output: pdf_document
---

### READING THE DATA ###

```{r, warning = F, message = F}
ds = read.csv("Joined4_NonsenseReplaced.csv")[, -c(1, 2)]

ds7 = read.csv("Joined7_NonsenseReplaced.csv")
ds$CANCER = ds7[,22]
ds$DIABETES = ds7[,21]
ds$I10_NPR = ds7$I10_NPR

d7 = read.csv("Dataset_Replaced.csv")
library(corrplot)
library(ROCR)
library(MASS)
library(tree)
library(pROC)
library(ggplot2)
library(randomForest)
```

```{r, out.width = "50%"}
deaths = character(68)
deaths = as.factor(ifelse(ds$DIED == 0, "Alive", "Dead"))
table(deaths)
plot(deaths)
```

### Using Both Number of Procedures and Cancer + Diabetes

### GLM

```{r}
m1 = glm(DIED ~ AGE + FEMALE + RACE + ZIPINC_QRTL + HOSP_REGION + CANCER + DIABETES + I10_NPR, 
         family = binomial, data = d7)
summary(m1)
pred_1 = as.factor(ifelse(predict(m1, d7, type = "response") <= 0.5, "Alive", "Dead"))
names(pred_1) = NULL
table_1 = table(pred_1, deaths)
table(pred_1, deaths)
table(pred_1, deaths)/68
sum(table(pred_1, deaths)[c(1,4)])/68
roc1 = performance(prediction(predict(m1, d7, type ='response'), deaths), "tpr", "fpr")
plot(roc1, colorize = T, main="ROC Curve for Logistic Regression Model")
```

### LDA

```{r}
m2 = lda(DIED ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + CANCER + DIABETES + YEAR + ZIPINC_QRTL + I10_NPR,
         data = ds)
pred_2 = predict(m2, newdata = ds)$class
pred_2 = ifelse(pred_2 == 0, "Alive", "Dead")
table(pred_2, deaths)
table(pred_2, deaths)/68
```

### Tree / Random Forest

```{r}
m3 = tree(deaths ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + CANCER + DIABETES + YEAR + ZIPINC_QRTL + I10_NPR,
         data = ds)
summary(m3)
pred_3 = predict(m3, data = ds, type = "class")
table(pred_3, deaths)
table(pred_3, deaths)/68
set.seed(1004)
test_ind = sample(68, 20)
dsTRAIN = ds[-test_ind, ]
dsTEST = ds[test_ind, ]
m4 = randomForest(DIED ~ AGE + FEMALE + RACE + HOSP_REGION + CANCER + DIABETES + YEAR + ZIPINC_QRTL + I10_NPR,
         data = dsTRAIN, importance = T, mtry = 7)
pred_4 = predict(m4, newdata = ds, type = "class")
pred_4[which(pred_4 >= .50)] = "DEAD" 
pred_4[which(pred_4 < .50)] = "ALIVE" 
table(pred_4, deaths)
table(pred_4, deaths)/68
```


### Using only number of procedures

### GLM

```{r}
m11 = glm(DIED ~ AGE + FEMALE + RACE + ZIPINC_QRTL + HOSP_REGION + COMOR + I10_NPR, 
         family = binomial, data = d7)
summary(m11)
pred_1 = as.factor(ifelse(predict(m11, d7, type = "response") <= 0.5, "Alive", "Dead"))
names(pred_1) = NULL
table_1 = table(pred_1, deaths)
table(pred_1, deaths)
table(pred_1, deaths)/68
sum(table(pred_1, deaths)[c(1,4)])/68
roc1 = performance(prediction(predict(m1, d7, type ='response'), deaths), "tpr", "fpr")
plot(roc1, colorize = T, main="ROC Curve for Logistic Regression Model")
```

### LDA

```{r}
m21 = lda(DIED ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + COMOR + I10_NPR,
         data = ds)
pred_2 = predict(m21, newdata = ds)$class
pred_2 = ifelse(pred_2 == 0, "Alive", "Dead")
table(pred_2, deaths)
table(pred_2, deaths)/68
```

### Tree / Random Forest

```{r}
m31 = tree(deaths ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + COMOR + I10_NPR,
         data = ds)
summary(m31)
pred_3 = predict(m31, data = ds, type = "class")
table(pred_3, deaths)
table(pred_3, deaths)/68
set.seed(1004)
test_ind = sample(68, 20)
dsTRAIN = ds[-test_ind, ]
dsTEST = ds[test_ind, ]
m41 = randomForest(DIED ~ AGE + FEMALE + RACE + HOSP_REGION + YEAR + ZIPINC_QRTL + COMOR + I10_NPR,
         data = dsTRAIN, importance = T, mtry = 7)
pred_4 = predict(m41, newdata = ds, type = "class")
pred_4[which(pred_4 >= .50)] = "DEAD" 
pred_4[which(pred_4 < .50)] = "ALIVE" 
table(pred_4, deaths)
table(pred_4, deaths)/68
```


## Using only COMOR replacement

### GLM

```{r}
m12 = glm(DIED ~ AGE + FEMALE + RACE + ZIPINC_QRTL + HOSP_REGION + CANCER + DIABETES, 
         family = binomial, data = d7)
summary(m11)
pred_1 = as.factor(ifelse(predict(m12, d7, type = "response") <= 0.5, "Alive", "Dead"))
names(pred_1) = NULL
table_1 = table(pred_1, deaths)
table(pred_1, deaths)
table(pred_1, deaths)/68
sum(table(pred_1, deaths)[c(1,4)])/68
roc1 = performance(prediction(predict(m1, d7, type ='response'), deaths), "tpr", "fpr")
plot(roc1, colorize = T, main="ROC Curve for Logistic Regression Model")
```

### LDA

```{r}
m22 = lda(DIED ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + CANCER + DIABETES,
         data = ds)
pred_2 = predict(m21, newdata = ds)$class
pred_2 = ifelse(pred_2 == 0, "Alive", "Dead")
table(pred_2, deaths)
table(pred_2, deaths)/68
```

### Tree / Random Forest

```{r}
m32 = tree(deaths ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + CANCER + DIABETES,
         data = ds)
summary(m32)
pred_3 = predict(m32, data = ds, type = "class")
table(pred_3, deaths)
table(pred_3, deaths)/68
set.seed(1004)
test_ind = sample(68, 20)
dsTRAIN = ds[-test_ind, ]
dsTEST = ds[test_ind, ]
m42 = randomForest(DIED ~ AGE + FEMALE + RACE + HOSP_REGION + YEAR + ZIPINC_QRTL + CANCER + DIABETES,
         data = dsTRAIN, importance = T, mtry = 7)
pred_4 = predict(m42, newdata = ds, type = "class")
pred_4[which(pred_4 >= .50)] = "DEAD" 
pred_4[which(pred_4 < .50)] = "ALIVE" 
table(pred_4, deaths)
table(pred_4, deaths)/68
```

```{r}
## Original models for later comparison
m10 = glm(DIED ~ AGE + FEMALE + RACE + ZIPINC_QRTL + HOSP_REGION + COMOR, 
         family = binomial, data = d7)
m20 = lda(DIED ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + CANCER + DIABETES,
         data = ds)
m30 = tree(deaths ~ AGE + FEMALE + RACE + HOSP_LOCTEACH + HOSP_REGION + YEAR + ZIPINC_QRTL + COMOR,
         data = ds)
m40 = randomForest(DIED ~ AGE + FEMALE + RACE + HOSP_REGION + YEAR + ZIPINC_QRTL + COMOR,
         data = dsTRAIN, importance = T, mtry = 7)

```

### Comparing the previous models to each other 

Remember that the order is original, using both changes, using just the number of procedures change, and then using just the COMOR change.

```{r}
summary(m10)
summary(m1)
summary(m11)
summary(m12)

summary(m20)
summary(m2)
summary(m21)
summary(m22)

summary(m30)
summary(m3)
summary(m31)
summary(m32)

importance(m40)
importance(m4)
importance(m41)
importance(m42)

varImpPlot(m40)
varImpPlot(m4)
varImpPlot(m41)
varImpPlot(m42)

```

